A Domain Specific Language for SpectreGuard

### 2020-12-11 paper/main.tex
Some final preparations

  A Spectre mitigation strategy in the form of a mechanism for marking variables
  as ``non-speculative'' was developed in the form of \textit{SpectreGuard}\cite{SpectreGuard}. In this paper, we
  develop an embedded domain specific language (EDSL) in Haskell to provide
  support for this feature, with a backend that generates C code. The static
  type system of Haskell is used so that many basic forms of copying secret
  variables to public variables are disallowed. Additionally, a static analysis
  was developed which will detect some cases of secret variables leaking to
  public variables via control flow constructs.

@@

\section{Introduction}
\label{sec:Introduction}
SpectreGuard is provides an abstraction that allows programmers to prevent
speculative access of certain variables~\cite{SpectreGuard}. As Spectre-based
attacks remain an issue, even in modern computer systems, such mitigations and
preventative measures are crucial to maintaining the security of software.  For
preserving program correctness, static analyses are desirable to prevent
programmers from accidentally successfully compiling code which copies a
``secret'' variable (a variable protected by SpectreGuard) to a ``public''
variable (a variable with no such protection). To this end, two interrelated
static analysis techniques are designed and implemented.

@@

\begin{itemize}
\item We describe a small domain specific language embedded in Haskell for representing
  simple C programs (Section~\ref{sec:Syntax}).
\item A code generator which generates C code from this EDSL is outlined
  (Section~\ref{sec:Syntax}).
\item We give a system of types which causes the expression language of this
  EDSL to be statically checked for variable copies that directly leak memory
  from secret variables to public variables (Section~\ref{sec:SensTypes}).
\item Another static analysis procedure is given which can find \textit{indirect}
  leak paths through control structures (Section~\ref{sec:Analysis}). This analysis
  procedure can be applied to EDSL code in the same way as the code generator
  can be applied to EDSL code. The results from this analysis take the form
  of a forest of multiway trees, providing information about these possible
  indirect leaks.
\item We provide several examples and describe how the system handles them
  (Section~\ref{sec:Examples}).
\end{itemize}

@@

Haskell is a purely functional programming language with strong static type
system based on System F and algebraic datatypes (ADTs)~\cite{HistoryOfHaskell}.
These features, particularly the type system, the algebraic datatypes and its
support for higher-order functions, make it a powerful tool for the creation of
EDSLs. Expressions, commands and programs (compound commands) can be
straightforwardly expressed using ADTs using a \textit{deep
embedding}~\cite{DeepAndShallow}.

A deep embedding directly exposes its abstract syntax tree (AST) as a value.
This allows EDSL programs to be easily manipulated and examined, such as for
interpretation, code generation and static analysis (the latter two we make use
of in this paper).

\section{Syntax}
\label{sec:Syntax}
A correspondence between the syntax of the EDSL, which is inherited from
Haskell, and the C language is given in Fig. \ref{fig:Syntax}. In the right column, an
overline indicates that the transformation from EDSL to C is recursively
performed to the expression or command underneath the overline.

This also outlines procedure that the C code generator uses. When code written
in the EDSL is transformed into C code, a name generator first annotates the
abstract syntax tree (AST) with globally unique names (enforcing the Barendregt
variable convention; see Chapter 2 in~\cite{BarendregtConversion}). This
annotated AST can be used by both the code generator and the information flow
leak analysis tool (the latter is described in Section \ref{sec:Analysis}). The
code generator performs a transformation from code in the syntax given in the
left column of Fig.~\ref{fig:Syntax} to the right column of the figure. This is
transformation recursively applies itself when it reaches the parts marked with
an overline in the right column of the figure. It is worth noting that, as a
result of this technique, the names in the generated C code will not generally
be the same as the names in the EDSL code.

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity type information at
runtime. Such a recovery is made possible by the use of singletons~\cite{Singletons}.
\ttt{for} loops provide a particular example of where Haskell's support for higher-order function
can be used inside an EDSL to implement commands that involve the creation of
new variables.

@@

\section{Sensitivity Types}
\label{sec:SensTypes}
The Haskell type system is leveraged to perform a basic static consistency check
of the sensitivities of variables, so that a variable does not get assigned to a
variable of a different sensitivity. This is done by making the \verb|Expr| type
constructor take a phantom type argument~\cite{Phantom}. The phantom type
argument is the \textit{sensitivity type} of that expression.  The logical
inference rules corresponding to this check are given in Fig.
\ref{fig:SensTypes}. Note that the EDSL inherits the rest of Haskell's type
system in addition to these rules. Only the aspects of the type system that is
particularly relevant to the EDSL is reflected in the figure, often specialized
to the particular scenario encountered in using sensitivity types in the EDSL.
An introduction to this style of inference rule system, with many examples and
applications in the study of programming languages, is given in
\cite{HarperFoundations}.

Observe that an two expressions with different sensitivity types cannot be
combined into a larger compound expression. However, commands (represented by
the \verb|Cmd| type constructor) often have expressions embedded in them and
commands can be sequenced with each other.  There is no restriction on how two
commands can be combined, even if they involve expressions with different
sensitivity types. This freedom in combining commands independently of
sensitivity foreshadows the essence of the problem we will encounter in
Section \ref{sec:Analysis}. For now, note that all control flow constructs are
part of the command type (as opposed to the expression type).

Singletons~\cite{Singletons} are used to preserve this type level at runtime, so that the
code generator can generate SpectreGuard annotations where appropriate. These annotations can
then be used by the SpectreGuard mitigation system~\cite{SpectreGuard}.

In Fig. \ref{fig:SensTypes}, it is assumed that $\tau$ and $\sigma$ are given by the
following grammar:
\begin{gather*}
  \sigma ::= \ttt{Public}\;|\;\ttt{Secret}\\
  \tau ::= \ttt{Int}\;|\;\ttt{Ptr}\;\tau\\
  \alpha ::= \ttt{()}\;|\;\ttt{Expr $\sigma$ $\tau$}
\end{gather*}

@@

\begin{figure}
  \centering
\resizebox{0.67\linewidth}{!}{
\begin{minipage}{\linewidth}
  \centering
  \openup 3mm
\begin{gather*}
  \inference{}{\ttt{Public :: Sensitivity}}\\
  \inference{}{\ttt{Secret :: Sensitivity}}\\
  \inference{n \in \mathbb{N}}{\ttt{$n$ :: Int}}\\
  \inference{n \in \mathbb{N}}{\ttt{$n$ :: Expr $\sigma$ Int}}\\
  \inference{%
    \Gamma \vdash \ttt{x :: Expr $\sigma$ Int}
    & \Gamma \vdash \ttt{y :: Expr $\sigma$ Int}}
    {\Gamma \vdash \ttt{x + y :: Expr $\sigma$ Int}}\\
  \inference{%
    \Gamma \vdash \ttt{x :: Expr $\sigma$ Int}
    & \Gamma \vdash \ttt{y :: Expr $\sigma$ Int}}
    {\Gamma \vdash \ttt{x <? y :: Expr $\sigma$ Bool}}\\
  \inference{%
    \Gamma \vdash \ttt{a :: Expr $\sigma$ (Ptr $\tau$)}
    & \Gamma \vdash \ttt{i :: Expr $\sigma$ Int}}
    {\Gamma \vdash \ttt{a ! i :: Expr $\sigma$ $\tau$}}\\
  \inference{\Gamma \vdash \ttt{n :: Int}}{\Gamma \vdash \ttt{allocPublic @$\tau$ n :: Cmd (Expr Public (Ptr $\tau$))}}\\
  \inference{\Gamma \vdash \ttt{n :: Int}}{\Gamma \vdash \ttt{allocSecret @$\tau$ n :: Cmd (Expr Secret (Ptr $\tau$))}}\\
  \inference{\Gamma \vdash \ttt{x :: $\tau$}}{\Gamma \vdash \ttt{decl x :: Cmd (Expr Public $\tau$)}}\\
  \inference{%
    \Gamma \vdash \ttt{x :: Expr $\sigma$ $\tau$}
    & \Gamma \vdash \ttt{y :: Expr $\sigma$ $\tau$}}
    {\Gamma \vdash \ttt{x .= y :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{cond :: Expr $\sigma$ Bool}
    \\ \Gamma \vdash \ttt{t :: Cmd $\alpha$}
    &\Gamma \vdash \ttt{f :: Cmd $\alpha$}}
    {\Gamma \vdash \ttt{ifThenElse cond t f :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{cond :: Expr $\sigma$ Bool}
    & \Gamma \vdash \ttt{body :: Cmd $\alpha$}}
    {\Gamma \vdash \ttt{while cond body :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{init :: Expr $\sigma$ $\tau$}\\
    \Gamma \vdash \ttt{fn :: Expr $\sigma$ $\tau$ $\rarr$ (Expr $\sigma$ Bool, Cmd (), Cmd ())}
    }{\Gamma \vdash \ttt{for init fn :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{c :: Cmd (Expr $\sigma$ $\tau$)}
    \\ \Gamma,\ttt{(x :: Expr $\sigma$ $\tau$)} \vdash \ttt{rest :: Cmd $\alpha$}
  }{%
    \Gamma \vdash \ttt{(\textbf{do} \{ x <- c; rest \}) :: Cmd $\alpha$}
  }\\
  \inference{%
    \Gamma \vdash \ttt{c :: Cmd (Expr $\sigma$ $\tau$)}
    \\ \Gamma \vdash \ttt{rest :: Cmd $\alpha$}
  }{%
    \Gamma \vdash \ttt{(\textbf{do} \{ c; rest \}) :: Cmd $\alpha$}
  }
\end{gather*}
\end{minipage}}
  \caption{Typing rules for sensitivity types}
\label{fig:SensTypes}
\end{figure}

@@

A couple of simple examples illustrate the basic typing rules:

@@

\begin{lstlisting}[language=haskell]
  -- This example fails to type check
example2 :: Cmd ()
example2 = do
  array1 <- allocSecret @Int 8
  array2 <- allocPublic @Int 8
  (*@\color{red}array2 .= array1@*)
\end{lstlisting}

\noindent \ttt{example2} fails to type check since \ttt{array1} and \ttt{array2} have
different sensitivity annotations, while the type of the \ttt{(.=)} operation
requires that both arguments have the same sensitivity annotation. This restriction can
be seen in the typing rule for \ttt{(.=)} in Fig. \ref{fig:SensTypes}.

@@

The next slightly more complex example exhibits an important limitation of this sensitivity type checking:

@@

\noindent Even though this code type checks without error, the entire contents of the secret variable \ttt{secretArray} is leaked
to the public variable \ttt{publicArray}. The leak occurs despite the fact that the two variables have different
sensitivity annotations. This is the
result of information being indirectly leaked from the secret variable to the
public variable via a control flow construct: in this case, the innermost \ttt{for} loop. Such an example has been studied from programming language perspective before; for instance
in~\cite{InfoFlowAnalysis}.

@@

In Section \ref{sec:Examples}, a significant limitation of the sensitivity type checking
was outlined. In this section we address this limitation using a static analysis technique.
Possible leaks from secret variables to public variables via control flow constructs will
be analyzed and the results of this analysis will be described in the form of
a forest of trees, given in DOT format (which can then be converted to other formats,
such as PDF)~\cite{DOT}. The C code generated from \ttt{example4} is shown in Fig. \ref{fig:GenC4}.
Note that the secret array is named \verb|_fs_0| and the public array is named \verb|_fs_1| by the
code generator. From the original \ttt{example4}, the information flow analyzer generates the
graph given in Fig. \ref{fig:Analysis4}. Secret variables are represented as nodes with
a rectangular shape and public variables are represented as nodes with an elliptic shape. An
arrow going from one node to another represents a possible information leak from the first
node to the second node due to a control flow construct.

@@

An alternate way to view the scope stack is as a stack of \textit{relations} on
on names. Each item of the stack is a relation such that two names are related
if and only if the left-hand name is a public variable declared in the corresponding
scope and the right-hand name is a secret variable in a conditional within the scope.

@@

Whenever a declaration is encountered in EDSL code, the corresponding variable
name is added to the scope stack with \ttt{scopeAddName}. Whenever a control
structure is entered (such as a \ttt{for} loop or an \ttt{if} statement),
\ttt{scopePush} is called with a list of all secret variables (if any) in the
condition of that control structure. When an assignment is encountered, it
uses \ttt{publicNonLocals} to see if there are any secret variables associated
to an enclosing control structure, outside of which the given name has been declared.

@@

A significant recent work also related to the use of domain specific languages
with strong, static type systems to control information flow for security
purposes is \textit{Liquid Information Flow Control} by Polikarpova et
al.~\cite{Lifty} As its title suggests, this paper also builds on the theory of information flow. In that work, a novel domain specific programming language is
developed which has a static type system that is geared specifically towards
expressing security policies at a type level. As this is done within a static
type system, these invariants are machine-verified at compile-time.  Not only
can these policy violations be detected, the system they develop is also able to
fix leaks.

More broadly, an example of a language being represented similarly, as an
pure expression language separated from a command language, can be found in
Winskel's 1993 book in the form of the \textbf{IMP} language.~\cite{WinskelBook} This
book provides an in-depth look at the three major mathematical formalizations
of programming language semantics (axiomatic semantics, operational semantics and
denotational semantics) and uses them to analyze \textbf{IMP}.

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{Singletons} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker~\cite{CertProg, DepHaskSpec}. A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique~\cite{DepartedProofs}.

Another possible, related, future development would be the to attempt to add a
mechanism which automatically fixes some (or all) leaks. Such a system has precendent
in the previously mentioned paper \textit{Liquid Information Flow Control}~\cite{Lifty}.
This could give programmers an additional tool towards ensuring correctness, with
respect to SpectreGuard annotations.

@@

Other possible leak paths should also be investigated. While outside the scope of this paper,
it there could be leaks which occur under circumstances other than direct copies or the indirect
leaks described in Section~\ref{sec:Analysis}. In future work, it would be useful to either
classify more leak paths (and deal with them appropriately) or prove that, in fact, no such additional
leak paths exist (at least, within a certain threat model).

@@

Building upon the SpectreGuard work~\cite{SpectreGuard}, we have constructed
a system which allows the programmer greater assurance that the SpectreGuard
mechanism will not be accidentally circumvented due to certain forms of
leaks from secret variables to public variables. This analysis checks for
both direct copies from a secret variable to a public variable, as well as
a certain class of leaks where control flow depends on a secret variable.

>>>>

  A Spectre mitigation strategy which allows variables to be marked
  as ``non-speculative'' was developed in the form of \textit{SpectreGuard}\cite{SpectreGuard}. In this paper, we
  develop an embedded domain specific language (EDSL) in Haskell to provide
  support for this feature, with a backend that generates C code. The static
  type system of Haskell is used so that basic forms of copying secret
  variables to public variables are disallowed. Additionally, a static analysis
  was developed which will detect some cases of secret variables leaking to
  public variables via control flow constructs.

@@

\section{Introduction}
\label{sec:Introduction}
SpectreGuard provides an abstraction that allows programmers to prevent
speculative access of certain variables~\cite{SpectreGuard}. As Spectre-based
attacks remain an issue, even in modern computer systems, such mitigations and
preventative measures are crucial to maintaining the security of software.  For
preserving program correctness with respect to SpectreGuard annotations, static analyses are desirable to prevent
programmers from accidentally compiling code which copies a
``secret'' variable (a variable protected by SpectreGuard) to a ``public''
variable (a variable with no such protection). To this end, two interrelated
static analysis techniques are designed and implemented: one within the Haskell type system and
one outside the type system.

@@

\begin{itemize}
\item We describe a small domain specific language embedded in Haskell for representing
  simple C programs (Section~\ref{sec:Syntax}).
\item A code generator which generates C code from this EDSL is outlined
  (Section~\ref{sec:Syntax}).
\item We give a system of types which causes the expression language of this
  EDSL to be statically checked for variable copies that directly leak memory
  from secret variables to public variables (Section~\ref{sec:SensTypes}).
\item Another static analysis procedure is given which can find \textit{indirect}
  leak paths through control structures (Section~\ref{sec:Analysis}). This analysis
  procedure can be applied to EDSL code in the same way as the code generator
  can be applied to EDSL code. The results from this analysis take the form
  of a forest of multiway trees, providing information to the programmer about these possible
  indirect leaks.
\item We provide several examples and describe how the system handles them
  (Section~\ref{sec:Examples}).
\end{itemize}

@@

Haskell is a purely functional programming language with a strong static type
system based on System F and algebraic datatypes (ADTs)~\cite{HistoryOfHaskell}.
These features, particularly the type system, the algebraic datatypes and its
support for higher-order functions, make it a powerful tool for the creation of
EDSLs. Expressions, commands and programs (compound commands) can be
straightforwardly expressed using ADTs to construct a \textit{deep
embedding}~\cite{DeepAndShallow}.

A deep embedding of an EDSL directly exposes its abstract syntax tree (AST) as a value.
This allows EDSL programs to be easily manipulated and examined, such as for the purposes of
interpretation, code generation and static analysis (the latter two we make use
of in this paper).

\section{Syntax}
\label{sec:Syntax}
A correspondence between the syntax of the EDSL, which is in part inherited from
Haskell, and the C language is given in Fig. \ref{fig:Syntax}. In the right column, an
overline indicates that the transformation from EDSL to C is recursively
performed to the expression or command underneath the overline.

This also outlines procedure that the C code generator uses. When code written
in the EDSL is transformed into C code, a name generator first annotates the
abstract syntax tree (AST) with globally unique names (enforcing the Barendregt
variable convention; see Chapter 2 in~\cite{BarendregtConversion}). This
annotated AST can be used by both the code generator and the information flow
leak analysis tool (the latter is described in Section \ref{sec:Analysis}). The
code generator performs a transformation from code in the syntax given in the
left column of Fig.~\ref{fig:Syntax} to the syntax in the right column of that figure. This
transformation recursively applies itself when it reaches the parts marked with
an overline in the right column of the figure. It is worth noting that, as a
result of the name generation technique used, the names in the generated C code will not generally
be the same as the names in the EDSL code.

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity information at
runtime. Such a recovery is made possible by the use of singletons~\cite{Singletons}.
\ttt{for} loops provide a particular example of where Haskell's support for higher-order functions
can be used inside an EDSL to implement commands that involve the creation of
new variables.

@@

\section{Sensitivity Types}
\label{sec:SensTypes}
The Haskell type system is leveraged to perform a basic static consistency check
of the sensitivities of variables, so that the contents of a variable does not get assigned to a
variable of a different sensitivity. This is done by making the \verb|Expr| type
constructor take a phantom type argument~\cite{Phantom} representing the sensitivity level. The phantom type
argument is the \textit{sensitivity level} of that expression.  The logical
inference rules corresponding to this check are given in Fig.
\ref{fig:SensTypes}. Note that the EDSL inherits the rest of Haskell's type
system in addition to these rules. Only the aspects of the type system that are
particularly relevant to the EDSL are reflected in the figure, often specialized
to the particular scenario encountered in using sensitivity levels in the EDSL.
An introduction to this style of inference rule system, with many examples and
applications in the study of programming languages, is given in
\cite{HarperFoundations}.

Observe that an two expressions with different sensitivity levels cannot be
combined into a larger compound expression. However, commands (represented by
the \verb|Cmd| type constructor) often have expressions embedded in them and
commands can be sequenced with each other.  There is no restriction on how two
commands can be combined, even if they involve expressions with different
sensitivity levels. This freedom in combining commands independently of
sensitivity foreshadows the essence of the problem we will encounter in
Section \ref{sec:Analysis}. For now, note that all control flow constructs are
part of the command type, as opposed to the expression type.

Singletons~\cite{Singletons} are used to preserve this type level information about sensitivities at runtime, so that the
code generator can generate SpectreGuard annotations where appropriate. These annotations can
then be used by the SpectreGuard mitigation system~\cite{SpectreGuard}.

In Fig. \ref{fig:SensTypes}, it is assumed that $\tau$ and $\sigma$ are metavariables given by the
following grammar:
\begin{gather*}
  \sigma ::= \ttt{Public}\;|\;\ttt{Secret}\\
  \tau ::= \ttt{Int}\;|\;\ttt{Ptr}\;\tau\\
  \alpha ::= \ttt{()}\;|\;\ttt{Expr $\sigma$ $\tau$}
\end{gather*}

@@

\begin{figure}
  \centering
\resizebox{0.67\linewidth}{!}{
\begin{minipage}{\linewidth}
  \centering
  \openup 3mm
\begin{gather*}
  \inference{}{\ttt{Public :: Sensitivity}}\\
  \inference{}{\ttt{Secret :: Sensitivity}}\\
  \inference{n \in \mathbb{N}}{\ttt{$n$ :: Int}}\\
  \inference{n \in \mathbb{N}}{\ttt{$n$ :: Expr $\sigma$ Int}}\\
  \inference{%
    \Gamma \vdash \ttt{x :: Expr $\sigma$ Int}
    & \Gamma \vdash \ttt{y :: Expr $\sigma$ Int}}
    {\Gamma \vdash \ttt{x + y :: Expr $\sigma$ Int}}\\
  \inference{%
    \Gamma \vdash \ttt{x :: Expr $\sigma$ Int}
    & \Gamma \vdash \ttt{y :: Expr $\sigma$ Int}}
    {\Gamma \vdash \ttt{x <? y :: Expr $\sigma$ Bool}}\\
  \inference{%
    \Gamma \vdash \ttt{a :: Expr $\sigma$ (Ptr $\tau$)}
    & \Gamma \vdash \ttt{i :: Expr $\sigma$ Int}}
    {\Gamma \vdash \ttt{a ! i :: Expr $\sigma$ $\tau$}}\\
  \inference{\Gamma \vdash \ttt{n :: Int}}{\Gamma \vdash \ttt{allocPublic @$\tau$ n :: Cmd (Expr Public (Ptr $\tau$))}}\\
  \inference{\Gamma \vdash \ttt{n :: Int}}{\Gamma \vdash \ttt{allocSecret @$\tau$ n :: Cmd (Expr Secret (Ptr $\tau$))}}\\
  \inference{\Gamma \vdash \ttt{x :: $\tau$}}{\Gamma \vdash \ttt{decl x :: Cmd (Expr Public $\tau$)}}\\
  \inference{%
    \Gamma \vdash \ttt{x :: Expr $\sigma$ $\tau$}
    & \Gamma \vdash \ttt{y :: Expr $\sigma$ $\tau$}}
    {\Gamma \vdash \ttt{x .= y :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{cond :: Expr $\sigma$ Bool}
    \\ \Gamma \vdash \ttt{t :: Cmd $\alpha$}
    &\Gamma \vdash \ttt{f :: Cmd $\alpha$}}
    {\Gamma \vdash \ttt{ifThenElse cond t f :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{cond :: Expr $\sigma$ Bool}
    & \Gamma \vdash \ttt{body :: Cmd $\alpha$}}
    {\Gamma \vdash \ttt{while cond body :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{init :: Expr $\sigma$ $\tau$}\\
    \Gamma \vdash \ttt{fn :: Expr $\sigma$ $\tau$ $\rarr$ (Expr $\sigma$ Bool, Cmd (), Cmd ())}
    }{\Gamma \vdash \ttt{for init fn :: Cmd ()}}\\
  \inference{%
    \Gamma \vdash \ttt{c :: Cmd (Expr $\sigma$ $\tau$)}
    \\ \Gamma,\ttt{(x :: Expr $\sigma$ $\tau$)} \vdash \ttt{rest :: Cmd $\alpha$}
  }{%
    \Gamma \vdash \ttt{(\textbf{do} \{ x <- c; rest \}) :: Cmd $\alpha$}
  }\\
  \inference{%
    \Gamma \vdash \ttt{c :: Cmd (Expr $\sigma$ $\tau$)}
    \\ \Gamma \vdash \ttt{rest :: Cmd $\alpha$}
  }{%
    \Gamma \vdash \ttt{(\textbf{do} \{ c; rest \}) :: Cmd $\alpha$}
  }
\end{gather*}
\end{minipage}}
  \caption{Typing rules for sensitivity levels}
\label{fig:SensTypes}
\end{figure}

@@

We now give a couple of simple examples illustrate the basic typing rules.

@@

\pagebreak
\begin{lstlisting}[language=haskell]
  -- This example fails to type check
example2 :: Cmd ()
example2 = do
  array1 <- allocSecret @Int 8
  array2 <- allocPublic @Int 8
  (*@\color{red}array2 .= array1@*)
\end{lstlisting}

\noindent \ttt{example2} fails to type check since \ttt{array1} and \ttt{array2} have
different sensitivity levels, while the type of the \ttt{(.=)} operation
requires that both arguments have the same sensitivity level. This restriction can
be seen in the typing rule for \ttt{(.=)} in Fig. \ref{fig:SensTypes}.

@@

The next slightly more complex example exhibits an important limitation of this sensitivity level checking:

@@

\noindent Even though this code type checks without error, the entire contents of the secret variable \ttt{secretArray} is leaked
to the public variable \ttt{publicArray}. The leak occurs despite the fact that the two variables have different
sensitivity annotations. This is the
result of information being indirectly leaked from the secret variable to the
public variable via a control flow construct: in this case, the innermost \ttt{for} loop. Such an example has been studied from programming language perspective in the
context of computer security before, for instance
in~\cite{InfoFlowAnalysis}.

@@

In Section \ref{sec:Examples}, a significant limitation of the sensitivity level type checking
was outlined. In this section we address this limitation using a static analysis technique.
Possible leaks from secret variables to public variables via control flow constructs will
be analyzed and the results of this analysis will be described in the form of
a forest of trees, given in DOT format (which can then be converted to other formats,
such as PDF)~\cite{DOT}. The C code generated from \ttt{example4} is shown in Fig. \ref{fig:GenC4}.
Note that the secret array is named \verb|_fs_0| and the public array is named \verb|_fs_1| by the
code generator. From the original \ttt{example4}, the information flow analyzer generates the
graph given in Fig. \ref{fig:Analysis4}. Secret variables are represented as nodes with
a rectangular shape and public variables are represented as nodes with an elliptic shape. An
arrow going from one node to another represents a possible information leak from the first
node to the second node due to a control flow construct.

@@

An alternate way to view the scope stack is as a stack of \textit{relations} on
on names. Each item of the stack is a relation such that two names are related
if and only if the left-hand name is a public variable declared in the corresponding
scope and the right-hand name is a secret variable in a conditional immediately within the scope.

@@

Whenever a declaration is encountered in EDSL code, the corresponding variable
name is added to the scope stack frame with \ttt{scopeAddName}. Whenever a control
structure is entered (such as a \ttt{for} loop or an \ttt{if} statement),
\ttt{scopePush} is called with a list of all secret variables (if any) in the
condition of that control structure. When an assignment is encountered, it
uses \ttt{publicNonLocals} to see if there are any secret variables associated
to an enclosing control structure, outside of which the given name has been declared.

@@

A significant recent work also related to the use of domain specific languages
with strong, static type systems to control information flow for security
purposes is \textit{Liquid Information Flow Control} by Polikarpova et
al.~\cite{Lifty} As its title suggests, this paper also builds on the theory of information flow. In that work, a novel domain specific programming language is
developed which has a static type system that is geared specifically towards
expressing security policies at a type level. As this is done within a static
type system, these invariants are machine-verified at compile-time.  Not only
can policy violations be detected and reported, the system they develop is also able to
fix leaks automatically.

More broadly, an example of a language being represented similarly, as an
pure expression language separated from a command language, can be found in
Winskel's 1993 book in the form of the \textbf{IMP} language. This
book provides an in-depth look at the three major mathematical formalizations
of programming language semantics (axiomatic semantics, operational semantics and
denotational semantics) and uses them to analyze \textbf{IMP}.~\cite{WinskelBook}

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{Singletons} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker~\cite{CertProg, DepHaskSpec}. A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique~\cite{DepartedProofs}. Another
way to expand the expressiveness of the type system is through the use of \textit{refinement
types}~\cite{RefinementTypes}, which have a Haskell implementation in the form of Liquid Haskell~\cite{LiquidHaskell}.
Refinement types allow types which are given by restricting existing types using a predicate that
the values of the type must satisfy. The predicate is written as a statement in
a decidable logical framework.

Another possible, related, future development would be the to attempt to add a
mechanism which automatically fixes some (or all) leaks. Such a system has precendent
in the previously mentioned paper \textit{Liquid Information Flow Control}. This work uses refinement types, of the sort mentioned previously.~\cite{Lifty}
This could give programmers an additional tool towards ensuring correctness, with
respect to SpectreGuard annotations.

@@

Other possible leak paths should also be investigated. While outside the scope of this paper,
there could be leaks which occur under circumstances other than direct copies or the indirect
leaks described in Section~\ref{sec:Analysis}. In future work, it would be useful to either
classify more leak paths (and deal with them appropriately) or prove that, in fact, no such additional
leak paths exist (within a certain threat model).

@@

Building upon the SpectreGuard work~\cite{SpectreGuard}, we have constructed
a system which allows the programmer greater assurance that the SpectreGuard
mechanism will not be accidentally circumvented due to certain forms of
leaks from secret variables to public variables. This analysis checks for
both direct copies from a secret variable to a public variable, as well as
a certain class of leaks that can occur when control flow depends on a secret variable.
